---
output:
  github_document:
    html_preview: true
---

```{r, include=FALSE}
library(tidyverse)
library(pander)
```

# SOC 371: Criminology

### Summer 2020

Instructor: [Charles Lanfear](mailto:clanfear@uw.edu)
Office: 255 Savery Hall

## Class Meetings 

**Lecture Zoom**: Monday / Wednesday 9:40-11:50

**Office Hours**: Tuesday: 12:00 - 1:30

## Overview

This course provides an introduction to sociological criminology. We begin by defining crime and law, then surveying criminology's classical foundations. The majority of the course will then be spent studying modern sociological theories of crime. This course emphasizes the underlying causal models of these theories, corresponding assumptions, and levels of explanation--structural, neighborhood, group, and individual. We will study these theories mainly with primary source material--the words of the original theorists and later contributors. We will also relate these theories to real-world social problems and policies, such as policing, mass incarceration, gun violence, genocide, and segregation.

### Course Objectives

* Provide you with an understanding of basic theories, concepts, and methods used by criminologists.

* Show you how criminological theories apply to real-world phenomena.

* Make you a critical consumer of claims about crime and crime policies in the news, social media, and political campaigns.

* Develop your critical and analytical skills through discussion and written assignments.

### [Course Website](https://clanfear.github.io/SOC371/)
[clanfear.github.io/SOC371/](https://clanfear.github.io/SOC371/)

The course website is accessible without a UWnetID and features all of the slides, links to readings, and assignments. Note many readings are, however, pay-walled and require a UWnetID to access. [See this page for instructions for adding a UW Library proxy to your browser for off-campus access to articles.](https://www.lib.washington.edu/help/connect/tools)

### [Course Canvas](https://canvas.uw.edu/courses/1382874)

The course Canvas will be used to access some readings, post discussion questions, and submit assignments. Do not use Canvas messages for communication with the instructor--it will not be monitored.

## Course Requirements

Grades will be assessed with the following breakdown:

| Item                   | Number | Percent of Grade |
|:----------------------:|:------:|:----------------:|
| Course Participation   | 18     |     10           |
| Discussion Questions   | 16     |     20           |
| Midterm Exams          |  2     |     40           |
| Final Exam             |  1     |     30           |

A score of at least 95% will guarantee a 4.0 in the course. Every 2.5 percentage points below 95 
will correspond to a .1 difference in the final grade (e.g. 90/100 is a 3.8).

#### Participation

Students are expected to complete readings and engage in discussion. Given the disruption presented by COVID-19, alternate evaluation (e.g. a 1-page reading and discussion response) is possible for students unable to attend a given synchronous lecture. Students must contact the instructor prior to absences to be considered for alternate evaluation.

#### Discussion Questions

Students will be required to submit at least two discussion questions via Canvas prior to each lecture meeting. These questions will be used to facilitate post-lecture discussion. Questions should be open-ended and reflect engagement with the material--that is, they delve deeper into ideas presented in the reading or raise critical questions. Questions are due by 6 PM the day before (i.e. 6 PM Sunday for Monday lecture's readings). If there are multiple readings, each question should address a different reading. 

#### Exams

There will be two midterm and one final exam. Exams will focus on application of concepts rather than regurgitation of knowledge. Consequently, they are open book and you will have at least a day to work on them. The format will consist of a number of short answer and essay questions. Students are expected to reference course readings with clear citations, e.g. "Shaw & McKay (1942) proposed delinquent subcultures as a mechanism for stability in neighborhood crime rates." Students may refer to other materials, but should do so sparingly and not the exclusion of relevant course material. If material from outside the course is cited, you must add a *Works Cited* section at the end of the exam with a full, properly formatted (e.g. ASA format) citation for the work.

## Course Texts and Readings

There is no textbook for this course. Rather, there are two required books documenting important studies in criminology--books you are likely to find on the shelf of a respectable criminologist. Both books are freely available as eBooks through the UW library and affordable in print.

* [Divergent Social Worlds (2010) Ruth D. Peterson & Lauren J. Krivo](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71175976510001451)

* [Darfur and the Crime of Genocide (2009) John Hagan & Wenona Rymond-Richmond](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71240618630001451) 

Other readings will be available via files posted on Canvas and links on the course website. These readings will primarily be academic journal articles. Note that journal articles are targeted at academic audiences. It is likely you will find them challenging to read and may not understand everything discussed--this is fine! Allot yourself more time for article reading than you would for a textbook chapter of comparable length. At the end of each lecture you will also receive a brief overview of what to expect and look for in the readings for the next lecture. If you find an article--or section of an article--particularly challenging or unclear, please contact the instructor, come to office hours to chat, or bring it up in discussion. 

## Content Warning

The content in this course engages with many emotionally and intellectually challenging topics, such as racism, sexism, violence, incarceration, and genocide. I will do my best to make the classroom environment a welcoming environment for confronting these topics. Do not hesitate to reach out if you have questions or concerns about encountering any course material. 

## Contacting the Instructor

You are encouraged to contact my via email with questions or concerns. Note, it may take up to 48 weekday hours to respond to your email.

## Course Outline

```{r, echo=FALSE, message=FALSE, warning=FALSE}
course_outline <- readxl::read_excel("../ignore/course_workbook.xlsx", sheet=1) %>% filter(Ignore!=1)
course_readings <- readxl::read_excel("../ignore/course_workbook.xlsx", sheet=2) %>% filter(Ignore!=1) %>%
  select(Lecture, Citation)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
course_outline %>%
  select(-Week, -Ignore, -Readings) %>%
  arrange(Date) %>%
  mutate(Date = format(Date, "%b %e")) %>%
  mutate_all(~ ifelse(is.na(.), "", .)) %>%
  pander()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
course_outline %>%
  select(Week, Lecture, Date, Topic) %>%
  filter(!is.na(Lecture)) %>%
  left_join(course_readings, by = "Lecture") %>%
  select(-Lecture, Readings = Citation) %>%
  group_by(Date) %>%
  mutate(reading_number = as.character(row_number())) %>%
  mutate(reading_number = ifelse(!is.na(lead(reading_number)) | !is.na(lag(reading_number)), paste0("(",reading_number,") "), "")) %>%
  summarize(Topic = first(Topic), Readings = paste(paste0(reading_number, Readings), collapse = "<br><br>")) %>%
  mutate(Readings = ifelse(str_detect(Readings, "NA"), "No assigned reading", Readings)) %>%
  mutate_at(vars(-Date), ~ ifelse(is.na(.), "", .)) %>%
  mutate(Topic = case_when(
    is.na(lag(Topic)) ~ Topic,
    lag(Topic) == Topic ~ paste0(Topic, " 2"),
    lead(Topic) == Topic ~ paste0(Topic, " 1"),
    TRUE ~ Topic
  )) %>%
  arrange(Date) %>%
  mutate(weekday = lubridate::wday(Date, label=TRUE)) %>%
  mutate(Date = paste0(weekday, ", ", format(Date, "%b %e"))) %>%
  mutate(`Topic and Readings` = paste0("**",Topic,"**", "<br>", Readings)) %>%
  select(Date, `Topic and Readings`) %>%
  pander()
```

## Accommodation Policy

Your experience in this class is important to me. The policies below cover accommodations for disability or religious activities. If you have other concerns 
related to participating in this class--such as difficulty with course materials or the pace of the course--
please contact me. I want this course to be maximally useful and accessible to all students.

### Religious Activities

Washington state law requires that UW develop a policy for accommodation of student absences or significant hardship due to reasons of faith or conscience, or for organized religious activities. Accommodations must be requested within the first two weeks of this course using the Religious Accommodations Request form available at:  [https://registrar.washington.edu/students/religious-accommodations-request/](https://registrar.washington.edu/students/religious-accommodations-request/)

### Disability Access  

It is the policy and practice of the University of Washington to create inclusive and accessible learning environments consistent with federal and state law. If you experience barriers based on disability, [please seek a meeting with DRS to discuss and address them](http://depts.washington.edu/uwdrs/). If you have already established accommodations with DRS, please communicate your approved accommodations to me at your earliest convenience so we can discuss your needs in this course. Disability Resources for Students (DRS) offers resources and coordinates reasonable accommodations for students with disabilities.  Reasonable accommodations are established through an interactive process between you, your instructor (me) and DRS. DRS information can be found at: [http://depts.washington.edu/uwdrs/](http://depts.washington.edu/uwdrs/)

## Late Work Policy

Late assignments will be penalized by 25% per day without official documentation of an unavoidable cause
for absence or inability to complete the assigngment on time. Valid unavoidable causes are limited to those
described in [Student Governance Policies Chapter 112, Subsection 1.B](http://www.washington.edu/admin/rules/policies/SGP/ScholRegCH112.html).

## Academic Misconduct Policy

As a University of Washington student, you are bound by the university's [student conduct policy](http://www.washington.edu/admin/rules/policies/SGP/SPCH209.html). Academic misconduct, including plagiarism--which includes copying any material from any source including fellow students without attribution--will be referred to the Community Standards & Student Conduct office. If you are not sure a particular practice is acceptable, please contact me via email or in office hours.

## Acknowledgements

This course is based, with permission, on Dr. Ross Matsueda's University of Washington SOC371 Criminology lecture.