---
output:
  github_document:
    html_preview: false
---

```{r, include=FALSE}
library(tidyverse)
library(pander)
```

# SOC 371
## Criminology
### University of Washington

Instructor: [Charles Lanfear](mailto:clanfear@uw.edu)
Office: 255 Savery Hall

## Class Meetings 

**Lecture Zoom**: Monday / Wednesday 9:40-11:50

**Office Hours**: Tuesday: 12:00 - 1:30

## Overview

This course provides an introduction to sociological criminology. We begin by defining crime and law, then surveying criminology's classical foundations. The majority of the course will then be spent studying modern sociological theories of crime. This course emphasizes the underlying causal models of these theories, corresponding assumptions, and levels of explanation--structural, neighborhood, group, and individual. We will study these theories with primary source material--the words of their originators and key contributors. We will also relate these theories to real-world social problems and policies, such as policing, mass incarceration, gun violence, genocide, and segregation.

### Course Objectives

* Provide you with an understanding of basic theories, concepts, and methods used by criminologists.

* Show you how criminological theories apply to real-world phenomena.

* Make you a critical consumer of media reports and politicians’ claims about crime and crime policies.

* Develop your critical and analytical skills through discussion and written assignments.

### [Course Website](https://clanfear.github.io/SOC371/)
[clanfear.github.io/SOC371/](https://clanfear.github.io/SOC371/)

The course website is accessible without a UWnetID and features all of the slides, links to readings, and assignments. Note many readings are pay-walled and require a UWnetID to access. [See this page for instructions for adding a UW Library proxy to your browser for off-campus access to articles.](https://www.lib.washington.edu/help/connect/tools)

### [Course Canvas](https://canvas.uw.edu/courses/1382874)

The course Canvas will be used to access some readings, post discussion questions, and submit assignments. Do not use Canvas messages for communication with the instructor--it will not be monitored for messages.

## Course Requirements

Grades will be assessed with the following breakdown:

| Item                   | Number | Percent of Grade |
|:----------------------:|:------:|:----------------:|
| Discussion Questions   | 16     |     10           |
| Midterm Exams          |  2     |     40           |
| Final Exam             |  1     |     30           |
| Essay                  |  1     |     20           |

A score of at least 95% will guarantee a 4.0 in the course. Every 2.5 percentage points below 95 
will correspond to a .1 difference in the final grade (e.g. 90/100 is a 3.8).

## Course Texts

There are two required texts for this course, both of which are freely available as eBooks through the UW library and affordable in print.

* [Divergent Social Worlds (2010) Ruth D. Peterson & Lauren J. Krivo](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71175976510001451)

* [Darfur and the Crime of Genocide (2009) John Hagan & Wenona Rymond-Richmond](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71240618630001451)

## Content Warning

The content in this course engages with many emotionally and intellectually challenging topics, such as racism, sexism, violence, incarceration, and genocide. I will do my best to make the classroom environment a welcoming environment for confronting these topics. Do not hesitate to reach out if you have questions or concerns about encountering any course material. 

## Contacting the Instructor

You are encouraged to contact my via email with questions or concerns. Note, it may take up to 48 weekday hours to respond to your email.

## Course Outline

```{r, echo=FALSE, message=FALSE, warning=FALSE}
course_outline <- readxl::read_excel("../ignore/course_workbook.xlsx", sheet=1) %>% filter(Ignore!=1)
course_readings <- readxl::read_excel("../ignore/course_workbook.xlsx", sheet=2) %>% filter(Ignore!=1) %>%
  select(Lecture, Citation)
course_outline %>%
  select(-Week, -Ignore, -Readings) %>%
  left_join(course_readings, by = "Lecture") %>%
  arrange(Date) %>%
  mutate(Date = format(Date, "%b %e")) %>%
  mutate_all(~ ifelse(is.na(.), "", .)) %>%
  pander()
```



## Late Work Policy

Late assignments will be penalized by 25% per day without official documentation of an unavoidable cause
for absence or inability to complete the assigngment on time. Valid unavoidable causes are limited to those
described in [Student Governance Policies Chapter 112, Subsection 1.B](http://www.washington.edu/admin/rules/policies/SGP/ScholRegCH112.html).

## Accomodation

Your experience in this class is important to me. If you have
already established accommodations with Disability Resources for Students (DRS), please
communicate your approved accommodations to me at your earliest convenience so we can
discuss your needs in this course. If you have non-disability related concerns 
for participating in this class--such as difficulty with course materials or the pace of the course--
please contact me. I want this course to be maximally useful and accessible to all students.

## Academic Misconduct

As a University of Washington student, you are bound by the university's [student conduct policy](http://www.washington.edu/admin/rules/policies/SGP/SPCH209.html). Academic misconduct, including plagiarism--which includes copying any material from any source including fellow students without attribution--will be referred to the Community Standards & Student Conduct office. If you are not sure a particular practice is acceptable, please contact me via email or in office hours.

## Acknowledgements

This course is based, with permission, on Dr. Ross Matsueda's University of Washington SOC371 Criminology lecture.