---
output:
  github_document:
    html_preview: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(pander)
```

# SOC 371: Criminology (Winter 2021)

## Contacts

* Instructor: [Chuck Lanfear (clanfear@uw.edu)](mailto:clanfear@uw.edu)
   * [Office Hours (Zoom): Tuesday: 11:00 - 12:00](https://washington.zoom.us/j/2979554878)
* Teaching Assistant: [Thiago Marques (thiagm@uw.edu)](mailto:thiagm@uw.edu)
   * [Office Hours (Zoom): TBA](https://washington.zoom.us/)

## Class Meetings 

* [**Lecture (Zoom)**: Monday / Wednesday 8:30-10:20](https://washington.zoom.us/j/97380045367)
* [**Section A (Zoom)**: Wednesday 3:30-4:20](https://washington.zoom.us/)
* [**Section B (Zoom)**: Wednesday 4:30-5:20](https://washington.zoom.us/)

## Overview

This course provides an introduction to sociological criminology. We begin by defining crime and law, then surveying criminology's classical foundations. The majority of the course will then be spent studying modern sociological theories of crime. This course emphasizes the underlying causal models of these theories, corresponding assumptions, and levels of explanation--structural, neighborhood, group, and individual. We will study these theories mainly with primary source material--the words of the original theorists and later contributors. We will also relate these theories to real-world social problems and policies, such as policing, mass incarceration, gun violence, genocide, and segregation.

### Course Objectives

* Provide you with an understanding of basic theories, concepts, and methods used by criminologists.

* Show you how criminological theories apply to real-world phenomena.

* Make you a critical consumer of claims about crime and crime policies in the news, social media, and political campaigns.

* Develop your critical and analytical skills through discussion and written assignments.

### [Course Website](https://clanfear.github.io/SOC371/)

[clanfear.github.io/SOC371/](https://clanfear.github.io/SOC371/)

The course website is accessible without a UWnetID and features all of the slides, links to readings, and assignments. Note many readings are, however, pay-walled and require a UWnetID to access. These can be found on the Canvas site or accessed via the UW Library. [See this page for instructions for adding a UW Library proxy to your browser for off-campus access to articles.](https://www.lib.washington.edu/help/connect/tools) Lecture videos will be posted within 36 hours of recording.

### [Course Canvas](https://canvas.uw.edu/courses/1436108)

The course Canvas will be used to access some readings, post discussion questions, and submit assignments. Do not use Canvas messages for communication with the instructor or TA--it will not be monitored.

## Course Requirements

Grades will be assessed with the following breakdown:

| Item                   | Number | Percent of Grade |
|:----------------------:|:------:|:----------------:|
| Course Participation   | 10     |     15           |
| Discussion Questions   | 17     |     15           |
| Midterm Exams          |  2     |     40           |
| Final Exam             |  1     |     30           |

A score of at least 95% will guarantee a 4.0 in the course. Every 2.0 percentage points below 95 (rounded *down*)
will correspond to a .1 difference in the final grade (e.g. 89% or 90% is a 3.7). You can calculate this yourself:

*Your Final Grade* = 4 - (95 - *Your Overall Percentage*) / 20

#### Participation

Students are expected to complete readings and engage in discussion during sections. Given the disruption presented by COVID-19, alternate evaluation is possible for students unable to attend one or more synchronous section meetings. Alternate evaluation takes the form of reading responses which *must not exceed 500 words* and should contain *for each reading*:

1. A basic summary of the reading
2. And one or both of:
   a. A personal reaction to the reading. That is, do you think the theory, its assumptions, and/or policy implications make sense or are missing important elements?
   b. A comparison and contrast with earlier readings and lecture content. You should make a statement about which you find more plausible and why. 

Please submit reading responses as a file attachment to the relevant participation assignment. Reading responses do not take the place of discussion questions. Students must still submit discussion questions. 

#### Discussion Questions

Students will be required to submit two discussion questions via Canvas prior to each lecture meeting. Please submit as text in the available entry boxes on Canvas. These questions will be used to facilitate section discussions (and occasionally lecture clarifications). Questions should be open-ended and reflect engagement with the material--that is, they delve deeper into ideas presented in the reading or raise critical questions. Questions are due by 5 PM the day before (i.e. 5 PM Sunday for Monday lecture's readings). If there are multiple readings, each question should address a different reading. Regardless of the number of readings, please always provide two questions.

#### Exams

There will be two midterm and one final exam. Exams will focus on application of concepts rather than regurgitation of knowledge. Consequently, they are open book and you will have a few days to work on them. The format will consist of a number of short answer and essay questions. Students are expected to reference course readings with clear citations, e.g. "Shaw & McKay (1942) proposed delinquent subcultures as a mechanism for stability in neighborhood crime rates." Students may refer to other materials, but should do so sparingly and not to the exclusion of relevant course material. If material from outside the course is cited, you must add a *Works Cited* section at the end of the exam with a full, properly formatted (e.g. ASA format) citation for the work.

Exams must be submitted to Canvas as a raw text (e.g. .txt) or common word processing document (e.g. .docx). Files that are "corrupt" or cannot be opened will be treated as handed in at the time a readable file is received. You may send the instructor a test file to open if you are unsure about file format compatibility. There are no minimum length requirements and maximums will be by question-specific word count, so don't waste your time playing with fonts, formatting, or margins. Exceeding the maximum word length for an answer will result in score deductions.

## Course Texts and Readings

There is no textbook for this course. Rather, there are two required books documenting important studies in criminology--books you are likely to find on the shelf of a respectable criminologist. Both books are freely available as eBooks through the UW library and affordable in print.

* [Divergent Social Worlds (2010) Ruth D. Peterson & Lauren J. Krivo](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71175976510001451)

* [Darfur and the Crime of Genocide (2009) John Hagan & Wenona Rymond-Richmond](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71240618630001451) 

Other readings will be available via files posted on Canvas and links on the course website. These readings will primarily be academic journal articles. Note that journal articles are targeted at academic audiences. It is likely you will find them challenging to read and may not understand everything discussed--this is fine! Allot yourself more time for article reading than you would for a textbook chapter of comparable length. At the end of each lecture you will also receive a brief overview of what to expect and look for in the readings for the next lecture. If you find an article--or section of an article--particularly challenging or unclear, please contact the teaching assistant, come to office hours to chat, or bring it up in discussion. 

## Content Warning

The content in this course engages with many emotionally and intellectually challenging topics, such as racism, sexism, violence including sexual violence, incarceration, and genocide. I will do my best to make the classroom environment a welcoming environment for confronting these topics. Do not hesitate to reach out if you have questions or concerns about encountering any course material. 

## Contacting the Instructor

You are encouraged to contact me via email with questions or concerns. Note, it may take up to 48 weekday hours to respond to your email.

## Course Outline

```{r, echo=FALSE, message=FALSE, warning=FALSE}
course_outline <- readxl::read_excel("../ignore/course_workbook.xlsx", sheet=1) %>% filter(Ignore!=1)
course_readings <- readxl::read_excel("../ignore/course_workbook.xlsx", sheet=2) %>% filter(Ignore!=1) %>%
  select(Lecture, Citation, Reading_Link)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
course_outline %>%
  select(Week, Lecture, Date, Topic, Link, Video) %>%
  filter(!is.na(Lecture)) %>%
  left_join(course_readings, by = "Lecture") %>%
  select(-Lecture) %>%
  mutate(Readings = paste0("[", Citation,"](", Reading_Link, ")")) %>%
  group_by(Date) %>%
  mutate(reading_number = as.character(row_number())) %>%
  mutate(reading_number = ifelse(!is.na(lead(reading_number)) | !is.na(lag(reading_number)), paste0("(",reading_number,") "), "")) %>%
  mutate(Link = ifelse(is.na(Link), "", Link)) %>%
  mutate(Topic = paste0("[", Topic, "](../", Link, ")")) %>%
  mutate(Video = ifelse(!is.na(Video), paste0("[(Zoom Recording)](", Video, ")"), "(No Recording Yet)")) %>%
  summarize(Topic = first(Topic), Readings = paste(paste0(reading_number, Readings), collapse = "<br><br>"), Video = first(Video)) %>%
  mutate(Readings = ifelse(str_detect(Readings, "NA"), "No assigned reading", Readings)) %>%
  mutate_at(vars(-Date), ~ ifelse(is.na(.), "", .)) %>%
  arrange(Date) %>%
  mutate(Date = format(Date, "%b %e")) %>%
  mutate(`Topic and Readings` = paste0("**",Topic,"**", "<br>", Video, "<br>", Readings)) %>%
  select(Date, `Topic and Readings`) %>%
  pander(split.tables=Inf)
```

## Accommodation Policy

Your experience in this class is important to me. The policies below cover accommodations for disability or religious activities. If you have other concerns 
related to participating in this class--such as difficulty with course materials or the pace of the course--
please contact me. I want this course to be maximally useful and accessible to all students.

### Religious Activities

Washington state law requires that UW develop a policy for accommodation of student absences or significant hardship due to reasons of faith or conscience, or for organized religious activities. Accommodations must be requested within the first two weeks of this course using the Religious Accommodations Request form available at:  [https://registrar.washington.edu/students/religious-accommodations-request/](https://registrar.washington.edu/students/religious-accommodations-request/)

### Disability Access  

It is the policy and practice of the University of Washington to create inclusive and accessible learning environments consistent with federal and state law. If you experience barriers based on disability, [please seek a meeting with DRS to discuss and address them](http://depts.washington.edu/uwdrs/). If you have already established accommodations with DRS, please communicate your approved accommodations to me at your earliest convenience so we can discuss your needs in this course. Disability Resources for Students (DRS) offers resources and coordinates reasonable accommodations for students with disabilities.  Reasonable accommodations are established through an interactive process between you, your instructor (me) and DRS. DRS information can be found at: [http://depts.washington.edu/uwdrs/](http://depts.washington.edu/uwdrs/)

## Late Work Policy

Late assignments will be penalized by 25% per day without official documentation of an unavoidable cause
for absence or inability to complete the assignment on time. Valid unavoidable causes are limited to those
described in [Student Governance Policies Chapter 112, Subsection 1.B](http://www.washington.edu/admin/rules/policies/SGP/ScholRegCH112.html). Discussion questions are not eligible for late turn-in (without documentation of unavoidable cause) and will receive a zero if submitted after the due time without prior approval.

## Academic Misconduct Policy

As a University of Washington student, you are bound by the university's [student conduct policy](http://www.washington.edu/admin/rules/policies/SGP/SPCH209.html). Academic misconduct, including plagiarism--which includes copying any material from any source including fellow students without attribution--will be referred to the Community Standards & Student Conduct office. If you are not sure a particular practice is acceptable, please contact me via email or in office hours.

## Acknowledgements

This course is based, with permission, on Dr. Ross Matsueda's University of Washington SOC371 Criminology lecture. It is also draws on material from Kate O'Neill's version of this same course and Ross Matsueda's SOC517 Deviance and Social control seminar.