---
title: "Introduction to Criminology"
subtitle: "SOC371, Winter 2021"
author: "Chuck Lanfear"
date: "Jan 4, 2020<br>Updated: `r gsub(' 0', ' ', format(Sys.Date(), format='%b %d, %Y'))`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ../style/xaringan-themer.css
    nature:
      highlightStyle: tomorrow-night-bright
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["center","top"]
---

# What is Criminology?

Sutherland: "The study of making of laws, the breaking of laws, and reacting to the breaking of laws."

--

* Making Laws: *Sociology of Law*

* Breaking Laws: *Criminology*

* Reacting to Law Breaking: *Sociology of Criminal Justice*

???

* While these all fall under broader criminology, most of the discipline focuses on violations of the law

* These are interrelated though: laws determine what counts as crime and also the response

* Assumptions about causes of crime shape reactions to breaking the law and legality of things in first place

* Closely related disciplines like LSJ and penology or criminal justice cover the other ends--we have courses elsewhere covering that.

--

&nbsp;

.center[*Criminology is a broad interdisciplinary field...*]

???

* Crim includes sociologists, psychologists, economists, philosophers, biologists, geographers, political scientists, legal scholars, and many others.

--

.center[*... but we're focusing on **sociological** criminology.*]

???

* But, criminology is dominated by sociological approaches and this is a sociological crim class, so that will be the focus.

* We'll be focusing on core mainstream theoretical frameworks and their foundations, which are largely dead white European guys.

* If you want to go deeper, more critical, or further afield, we can give you suggestions for further reading.

---

# Who am I?

Chuck Lanfear

* Instructor (not professor)

* Sociology PhD student

* My Office Hours (Zoom): Tuesday, 11:00 - 12:00

      * Thiago's (TA) Office Hours: Thursday, 4:00 - 5:00

--

My research:

* Neighborhood Social Control

* Criminal Opportunity and the Built Environment

* Racial Bias in Policing

???

Class is weighted a bit toward my areas of expertise. If you are aware of areas you're interested in that won't be covered, let us know early and I'll see if I can modify things.

---

# Objectives of Course

* Provide you with an understanding of basic theories, concepts, and methods used by criminologists.

   + Theories underlie policies

???

Theories are important. Crim theory can have big impacts, both good and bad. Broken windows--a theory--led to stop and frisk and order maintenance in the 1990s.

--

* Show you how criminological theories apply to real-world phenomena and current events.

   + How do we control crime?
   + How do we control police misconduct?
   + How do we address crime by nation-states?

--

* Make you a critical consumer of claims about crime and crime policies in the news, social media, and political campaigns.


--

* Develop your critical and analytical skills through discussion and written exams.


---
class: inverse
# Course Structure


---
# Two Components

* Lecture (Me)

   * Overviews of theories
   * Extensions and applications
   * Time to ask questions
   * Attendance appreciated but not required

--

* Section (Thiago)

   * Time to discuss topics
       * *You* will provide discussion questions
   * There *may* be activities
   * Participation required<sup>1</sup>

.footnote[[1] See the syllabus for an alternative to section participation.]

---
# Content

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
course_outline <- readxl::read_excel("../../ignore/course_workbook.xlsx", sheet=1) %>% filter(Ignore!=1)
course_readings <- readxl::read_excel("../../ignore/course_workbook.xlsx", sheet=2) %>% filter(Ignore!=1) %>%
  select(Lecture, Citation)
```

.small[
```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}
course_outline %>%
  select(Week, Date, Topic) %>%
  arrange(Date) %>%
  mutate(date_orig = Date, Date = format(Date, "%a, %b %e")) %>%
  arrange(date_orig) %>%
  select(-date_orig) %>%
  gt::gt()
```
]

---
class: inverse
## Course Requirements

---
# Grading

| Item                   | Number | Percent of Grade |
|:----------------------:|:------:|:----------------:|
| Course Participation   | 10     |     15           |
| Discussion Questions   | 17     |     15           |
| Midterm Exams          |  2     |     40           |
| Final Exam             |  1     |     30           |

* 95% or higher is a 4.0  

* Each 2.0% corresponds to a .1 final grade difference 

   + *Your Final Grade* = $4 - \frac{95 - \text{Percentage}}{20}$
   
   + e.g. 89/100 is a 3.7

   + This is rounded down<sup>1</sup>: 90/100 is also 3.7
   
.footnote[[1] I'll bump you up if you're *real close*.]

???

If distribution is looking low, will bump everyone up (not curve)

---
# Participation

Components:

* Active participation in section
   + Attendance $\neq$ Participation

* Demonstration of having read the material

--

If unable to attend, you can submit a 1-page (no more than 500 word) reading and discussion response *prior to scheduled section discussion*.

* This option must be preapproved by me/TA via email<sup>1</sup>

* Responses which are not preapproved will not be accepted

.footnote[[1] You can get approval for more than one at a time. Just let us know your situation.]

???

Contact me if you will always or usually be unable to attend section

---
# Discussion Questions

Submit two discussion questions via Canvas

   + Due 5 PM the day prior to relevant lecture
      + No late submissions or make-up
   + Address different readings unless there is only one
   + If there are more than two readings, pick just two for questions

--

Aim for open-ended and thought-provoking:

> Harcourt suggests that broken windows policing is based on extending the concept of harm to include disorderly conduct in public spaces. How does this relate to controlling language or writing that is perceived as harmful?

???

Questions should be open ended

Extensions or new applications for the material

Comparisons to other readings; e.g. X says this, but Y says this.

--

Avoid clarifications<sup>1</sup> or questions with clear objective answers:

"What is swiftness in classical theory and how is it related to crime today?"

.footnote[[1] You can *in addition* write clarification questions. These will be answered in section or lecture.]

???

Since these are meant to provoke discussion, they should be debatable, and not something directly empirically answerable. 




---
# Exams

**2 Midterms**

**1 Final**

Structure:
* Open-book with a few days to work
* Only short-answer questions
* Citations expected:
   + ex: "Shaw & McKay (1942) proposed delinquent subcultures as a mechanism for stability in neighborhood crime rates."
   + Material from *outside* course must have an entry in **Works Cited**
   + Assigned class material can just be author(s) and date
   + Any major format (e.g. ASA) acceptable but be consistent

--

*Emphasis of exams is demonstrating understanding, critical thinking, and application of concepts, not regurgitation of facts, names, or dates.*

---
## Course Texts and Readings

Two books:

* *Short* and *accessible* yet *major contributions* to criminology
* Available as eBooks through the UW library but affordable in print

--

[Ruth D. Peterson & Lauren J. Krivo (2010) *Divergent Social Worlds*](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71175976510001451)
  
   + An analysis of urban neighborhood crime from a structural theory of racial-spatial inequality.

--

[John Hagan & Wenona Rymond-Richmond (2009) *Darfur and the Crime of Genocide* ](https://alliance-primo.hosted.exlibrisgroup.com/permalink/f/kjtuig/CP71240618630001451) 
   
   + An analysis of the Darfur genocide from a differential social organization theory of collective racial violence.

--

If you need a supplementary reference text:

1. Kubrin, Stucky, & Krohn. (2009) *Researching Theories of Crime and Deviance.*
2. Crutchfield, Kubrin, Bridges & Weis. (2008) *Crime: Readings*


---
# The Approach

1. Context

   * Conditions of life
   * Responses and critiques

???

Theories are rooted in times and places

Academic work is a never-ending conversation--nothing exists in a vacuum

--

2. Theoretical Assumptions

   * Basis for theory
   * "If this is the case, then..."

???

All theories make some assumptions about how the world works

Based on these, they derive principles. A lot of science is about evaluating the assumptions and principles that follow from them.

Note theories based on wrong assumptions can sometimes make completely correct predictions, like heliocentric models of planets

--

3. Policy Implications

   * Causes imply interventions

???

Theories imply causes and causes imply things you can try to change

Interventions based on incorrect theories are more likely to have unexpected consequences--we'll talk about some later

--

4. Evidence

   * Quantitative
   * Qualitative

???

Reciprocal relationship between quant and qual

Qual typically generates theories and questions

Quant typically them and looks at generalizability

Qual also examines mechanisms and assumptions

---
class: inverse
# Some Metatheory

---
# Causal Models

All theories either imply or explicitly state a causal model.

--

Theories consist of:

1. Assumptions: Taken for granted statements

   + *The foundations of theories*
   
--

2. Propositions: Testable hypotheses or predictions of interest

   + *These typically emerge from assumptions*

--

In criminology, the key assumptions are often about human nature...

+ e.g. *Humans are rational utility maximizers* (Homo economicus)

???

Sometimes we purposely use assumptions we know aren't completely correct because the consequences are minor in our application. Perfectly rational actors are an example.

--

...or mechanisms...

+ e.g. *Disorder signals that a neighborhood is out of control.*

Assumptions are often fiercely contested and sometimes untestable.

---
# Causal Graphs

Scientists often use **causal graphs** to depict models.

<br>

.image-75[
![](img/x_causes_y.svg)
]

---
# Key Issues


* Reverse Causality

   + $Y$ may cause $X$ instead
   + We typically rely on *assumptions* about directionality

--

* Spuriousness

   + $X$ falsely appears to cause $Y$
   + Instead $X$ and $Y$ are both caused by another factor (e.g. $Z$)

--

* Mediation

  + Complete mediation means $X$ *only* effects $Y$ through $Z$.
  + Partial mediation means $X$ effects $Y$ directly *and* through $Z$.
  
--

* Moderation

  + $X$ causes $Y$, but by how much depends on $Z$.

---
# More Like Ice *Crime*

<br>

.image-75[
![](img/ice_cream_crime.svg)
]

.centernote[*Ice cream sales cause violent crime*]

---
# Spuriousness

<br>

.image-75[
![](img/hot_weather_crime.svg)
]

.centernote[*The effect of ice cream sales is **spurious**; hot weather causes both crime and ice cream sales.*]

???

When a paper says they control for something, this is what they're adjusting for.

---
# Mediation

<br>

.image-75[
![](img/people_outside.svg)
]

.centernote[*Hot weather causes crime only because more people are outside; the number of people outside **mediates** the effect of hot weather on crime*]

---
# Moderation

<br>

.image-75[
![](img/pandemic.svg)
]

.centernote[*Social distancing **moderates** these relationships, because now fewer people are outside on hot days and those people stay away from each other.*]

---
# It Gets Complicated

<br>

![](img/ce_system.svg)


---
# Using Graphs

Many of the theories in this course will be depicted with causal graphs.

--

All claims (propositions) about the world imply a theory (with assumptions).

--

When you encounter new claims, it can be clarifying to sketch the implied causal graph.

--

If the causal relationships don't make sense, it may reveal a flawed theory:

* Outcomes which happen before their causes (reverse causality)

* Obvious causes left out (spuriousness)

* Ambiguous or nonsensical mechanisms

* Unwarranted assumptions

--

.centernote[*Be critical consumers of theories*]

---
class:inverse
# Questions

---
# For Next Time

### Readings: 

* Trevino (1996) "Cesare Beccaria: Legal Reformer" in *The Sociology of Law*

* Radzinowicz (1966) "The Liberal Position" in *Ideology and Crime*

### Things to pay attention to:

1. The context the classical school emerged from

2. The assumptions of the classical school--human nature, the role of government
   + The assumed causes of crime
   + The proposed ideal responses which emerge from those assumptions